---
title: "AP_histo_size"
author: "Chloé Gilligan"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(dplyr)
if (!require(gridExtra)) {
  install.packages("gridExtra")
  library(gridExtra)
}

```

#Reading in data
```{r}
Size <- read.csv("../data/AST_HF_histo_size.csv", header = T, na.strings = c("", "NA")) %>% 
  select(full_sample_id, date, sex, stage, length, width) 
  #filter(treatment %in% c("Ambient", "Heat"))
```


#Calculating gemoetric mena and volume
```{r}
Size_calc <- Size %>%
  mutate(
    geo.mean = sqrt(length * width),
    geo.mean.vol = pi * ((geo.mean / 2) ^ 2)
  )
```

#Getting rid of weird outliers
```{r}
Size_calc <- Size_calc %>%
  filter(!(sex == "male" & stage == 3 & date == "7/5/22" & geo.mean > 150))
# Remove the outliers
Size_calc <- Size_calc %>%
  filter(!(sex == "male" & stage == 3 & date == "7/5/22" & geo.mean > 150))

```


# Define standard error function
```{r}
std.error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}
```

#Fixing dates
```{r}
unique_dates <- unique(Size_calc$date)
print(unique_dates)
print(length(unique_dates))
```



# Calculate means and standard errors
```{r}
# Convert date column to factor with desired labels
size.means <- Size_calc %>%
  group_by(date, sex, stage) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    date = factor(date, labels = c("7/5/22","7/14/21","7/20/22","7/27/22","8/3/22","8/9/22","8/24/22")))

#Getting rid of weid outlier
# Remove the sixth row
size.means <- size.means[-6, ]


# Ensure sex is treated as a factor
size.means$sex <- factor(size.means$sex, levels = c("male", "female"))
size.means$stage <- factor(size.means$stage)  # New line to treat stage as a factor 

```



# Set factor levels for date and sex
```{r}
#size.means$treatment <- factor(size.means$treatment, levels = c("field", "ambient", "high"))
#size.means$date <- factor(size.means$date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

#size.means$date <- factor(size.means$date, levels = c("2/25/21","4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

```

```{r}
#size.means <- size.means %>%
  #mutate(
    #treatment = factor(treatment, levels = c("Field", "Ambient", "High")),
    #date = factor(date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")),
    #sex = factor(sex, levels = c("male", "female"))
  #)

#GETTING RID OF FOR ONLY A TEST
```

# Filter data for females
```{r}
#size.means_female <- size.means %>% filter(sex == "female")

#size.means_female <- size.means %>% filter(sex == "female", treatment %in% c("Ambient", "Heat"))
#size.means_male <- size.means %>% filter(sex == "male", treatment %in% c("Ambient", "Heat"))

size.means_female <- size.means %>% filter(sex == "female")
size.means_male <- size.means %>% filter(sex == "male")

```


#Calculating volume means and stard errors
```{r}
#vol.means <- Size %>%
  #group_by( date, month, sex, treatment) %>%
  #summarise(
    #gm = mean(geo.mean.vol, na.rm = TRUE), 
    #sem = std.error(geo.mean.vol),
    #n = n())

#vol.means[vol.means == "NaN"] <- NA
#vol.means$month <- factor(vol.means$month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))

# vol.means$mm3 <- vol.means$gm / 10^9
```


#Volume
```{r}
#vol.means <- Size_calc %>%
 # group_by(date, sex, treatment) %>%
 # summarise(
  #  gm = mean(geo.mean.vol, na.rm = TRUE), 
   # sem = std.error(geo.mean.vol),
   # n = n(),
   # .groups = 'drop'
#  )

#vol.means <- vol.means %>%
#  mutate(
   # date = factor(date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
 # )

#vol.means <- Size_calc %>%
  #group_by(treatment, date, sex) %>%
  #summarise(gm = mean(geo.mean.vol, na.rm=TRUE), 
            #sem = std.error(geo.mean.vol, na.rm=TRUE),
            #n=n())


# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }
}

# Apply the custom function within summarise
```

#Volume for experimental script- Not working anymore
```{r}
#vol.means <- Size_calc %>%
 # group_by(treatment, date, sex) %>%
 # summarise(gm = mean(geo.mean.vol, na.rm = TRUE), 
          #  sem = std.error(geo.mean.vol),
          #  n = n())

#vol.means[vol.means == "NaN"] <- NA

#vol.means$mm3 <- vol.means$gm/10^9

#vol.means$date <- factor(vol.means$date, levels = c("2/25/21", "4/2/21", "6/2/21", "7/23/21", "8/25/21"))

# Check the filtered data
#print(size.means_filtered)

```
#Volume for experimental script
```{r}
vol.means <- Size_calc %>%
  group_by(date, sex, stage) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'  # This ensures the grouped output is not grouped further
  ) %>%
  mutate(
    gm = ifelse(is.nan(gm), NA, gm),
    sem = ifelse(is.nan(sem), NA, sem)
  )

# Ensure date column is treated as a factor with the correct order
vol.means$date <- factor(vol.means$date, levels = c("7/5/22","7/14/21","7/20/22","7/27/22","8/3/22","8/9/22","8/24/22"))

# Print the resulting dataframe
print(vol.means)

```


```{r}
# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }}

# Convert date column to Date format
Size_calc$date <- as.Date(Size_calc$date, format = "%m/%d/%y")

# Apply the custom function within summarise
vol.means <- Size_calc %>%
  group_by(date, sex) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'
  )

# Ensure that any "NaN" values are replaced with NA
#vol.means[vol.means == "NaN"] <- NA

# Convert date column to factor with desired levels
vol.means$date <- factor(vol.means$date, levels = c("2022-07-05", "2022-07-14", "2022-07-20", "2022-07-27", "2022-08-03", "2022-08-09", "2022-08-24"   ))
```

#Plotting test with stages
```{r}
plot_by_stage <- function(data, sex_label) {
  ggplot(data, aes(x = date, y = gm, color = stage)) +
    geom_point(size = 3) +
    labs(title = paste("Geometric Mean Diameter for", sex_label),
         x = "Months",
         y = "Geometric Mean Diameter",
         color = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(breaks = seq(0, max(data$gm, na.rm = TRUE), by = 10)) +
    scale_color_manual(values = c("1" = "#CCEEF9", "2" = "#A6E1F4", "3" = "#59C7EB", "4" = "#00A9E0", "5" = "#137899"))
}

plot_female <- plot_by_stage(size.means_female, "Females")
plot_male <- plot_by_stage(size.means_male, "Males")

grid.arrange(plot_female, plot_male, ncol = 2)

```



```{r}
plot_by_stage <- function(data, sex_label) {
  ggplot(data, aes(x = date, y = gm, color = stage)) +
    geom_point(size = 3) +
    labs(title = paste("Geometric Mean Diameter for", sex_label),
         x = "Months",
         y = "Geometric Mean Diameter",
         color = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(breaks = seq(0, max(data$gm, na.rm = TRUE), by = 10))
}

plot_female <- plot_by_stage(size.means_female, "Females")
plot_male <- plot_by_stage(size.means_male, "Males")

grid.arrange(plot_female, plot_male, ncol = 2)

```



#Plotting Males and Females
```{r}
# Plotting Females
plot_female <- ggplot(size.means_female, aes(x = date, y = gm)) +
  geom_point(size = 3, color = "blue") +
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(size.means_female$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Plotting Males
```{r}

# Plotting Males
plot_male <- ggplot(size.means_male, aes(x = date, y = gm)) +
  geom_point(size = 3, color = "red") +
  labs(title = "Geometric Mean Diameter for Males",
       x = "Months",
       y = "Geometric Mean Diameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(size.means_male$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing
```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```