---
title: "AST_histo_highfrequency"
author: "Chloé Gilligan"
date: "2024-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing ggpubr package
```{r}
install.packages("ggpubr")
```

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("RColorBrewer")
library(ggpubr)
```

## Reading in data and metadata
```{r}
library(readr)
data <- read_csv("../data/AST_HF_histo_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
select(full_sample_id, date, species, site, sex, StageI, StageII, StageIII, StageIV, StageV)
```


## Separating the "TP" from sample ID's the begin with AST, keeping them on the field samples
```{r}
library(stringr)
#data <- data %>%
  #mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 #gsub("-TP\\d+", "", full_sample_id), 
                                # full_sample_id))
#data <- data %>%
 # filter(!timepoint=="TP0")
```

#calculate the total number of gametes in a sample
```{r}

data <- data %>%
rowwise() %>%
mutate(total.gametes = sum(StageI,StageII,StageIII,StageIV,StageV,na.rm = T))

```

# assign a value of Stage 0 to samples with no gametes
#if = if total.gametes = 0, then add a value of 1 to a new column called "Stage0"
#else = if total.gamates > 0, then add a value of 0 to a new column called "Stage0"
```{r}
data <- data %>%
  mutate(Stage0 = ifelse(total.gametes==0, 1, 0))
```

#calculate relative abundance of stage by sex, timepoint, treatment
```{r}
Rel.Abundance <- data %>%
  group_by(sex, date) %>%
  summarise(
    total_gametes = sum(total.gametes, na.rm = TRUE),
    StageI = sum(StageI, na.rm = TRUE),
    StageII = sum(StageII, na.rm = TRUE),
    StageIII = sum(StageIII, na.rm = TRUE),
    StageIV = sum(StageIV, na.rm = TRUE),
    StageV = sum(StageV, na.rm = TRUE)
  ) %>%
  mutate(
    rel_abundance_StageI = StageI / total_gametes,
    rel_abundance_StageII = StageII / total_gametes,
    rel_abundance_StageIII = StageIII / total_gametes,
    rel_abundance_StageIV = StageIV / total_gametes,
    rel_abundance_StageV = StageV / total_gametes
  )

Rel.Abundance <- Rel.Abundance %>%
  filter(!sex=="NA") 
```

#Pivot longer for plotting
```{R}
Rel.Abundance_long <- Rel.Abundance %>%
  pivot_longer(cols = starts_with("rel_abundance_stage"), 
               names_to = "stage", 
               values_to = "value") %>%
  mutate(stage = factor(stage, 
levels = c("rel_abundance_StageI", "rel_abundance_StageII", "rel_abundance_StageIII", "rel_abundance_StageIV", "rel_abundance_StageV"), 
                        labels = c("Stage I", "Stage II", "Stage III", "Stage IV", "Stage V")))

```

# add lines of TP0 for both
#Creating a seperate dataset and plot for the field samples
```{r}
#Rel.Abundance_long_field <- Rel.Abundance_long %>%
 # filter(treatment=="Field")
 #mutate(date = as.Date(date, format = "%m/%d/%y"))  # Convert date column to Date type for proper ordering
```

```{r}
# Convert date column to Date type for proper ordering
Rel.Abundance_long$date <- as.Date(Rel.Abundance_long$date, format = "%m/%d/%y")

# Define the desired order of dates
desired_order <- as.Date(c("7/5/22", "7/14/22", "7/20/22", "7/27/22", "8/3/22", "8/9/22"), format = "%m/%d/%y")

# Reorder the levels of the date factor according to desired_order
Rel.Abundance_long$date <- factor(Rel.Abundance_long$date, levels = desired_order)

# Create the plot with reordered dates
ggplot(Rel.Abundance_long, aes(x = date, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~ sex) +
  theme_minimal() +
  labs(
    title = "Relative Abundance by Date",
    x = "Date",
    y = "Relative Abundance",
    color = "Stage"
  ) +
  scale_fill_manual(
    values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
    labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V')
  ) +
  theme_bw()

```





# Create the plot using date for x-axis labels
```{r}
ggplot(Rel.Abundance_long, aes(x = date, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~ sex) +
  theme_minimal() +
  labs(
    title = "Relative Abundance by Date",
    x = "Date",
    y = "Relative Abundance",
    color = "Stage"
  ) +
  scale_fill_manual(
    values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
    labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V')
  ) +
  #scale_x_date(date_labels = "%b %d, %Y", date_breaks = "4 day") +
  theme_bw()

```


