---
title: "AST_histo_abundance"
author: "Chloé Gilligan"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

This script will analyze gamete abundance in Astrangia poculata over time. The data is separated into 3 treatments (ambient temperature, high temperature, field). 


## Installing ggpubr package
```{r}
install.packages("ggpubr")
```

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("RColorBrewer")
library(ggpubr)
```

## Reading in data and metadata
```{r}
library(readr)
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
select(full_sample_id, timepoint, date, species, month, site,  sex, StageI, StageII, StageIII, StageIV, StageV, total_gametes)
metadata <- read_csv("../AST_histology_metadata.csv")
```


## Filtering out the TP from sample ID's the begin with AST, keeping them on the field samples
```{r}
library(stringr)
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))
```

## Merging data to include treatment and site from the metadata to the histology abundance and size data
```{r}
##merged_data <- merge(data, metadata, by = "full_sample_id")
##write_csv(merged_data, "../data/merged_data.csv")
```

## Select females
```{r}

## changing merged_data to data
Abundance_female <- data %>% 
  filter(sex == "female") 

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'date'
```

### Calculate relative proportion of each stage 
```{r}
Abundance_female <- Abundance_female %>%
  mutate(rel_prop_stage1 = (StageI / total_gametes) * 100,
         rel_prop_stage2 = (StageII / total_gametes) * 100, 
         rel_prop_stage3 = (StageIII / total_gametes) * 100, 
         rel_prop_stage4 = (StageIV / total_gametes) * 100,
         rel_prop_stage5 = (StageV / total_gametes) * 100)
```

### Change data from wide to long 
```{r}
Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>%
  select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
  na.omit()                    # Remove rows with NA in any other columns
```

### SUMMARIZE DATA
```{r}
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint) %>% 
  summarise(total_count = n())
```

### Exporting CSV of merged data
```{r}
# Specify the file path where you want to save the CSV file
file_path <- "merged.csv"

# Export the data frame to a CSV file
write.csv(data, file = file_path, row.names = FALSE)
file.exists(file_path)
```


## Plotting Data
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
  geom_bar(stat="identity") +
  #labs(x = NULL, y = "Relative Abundance", fill = "Stage") +
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
# scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() 
#Figure out if relative porportion or not, or if stage 0 is needed?
theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm"))



```




```{r}
library(ggplot2)
library(RColorBrewer)

# Basic plot to check data
basic_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity", position="fill", colour="black") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() 

# Display the basic plot
print(basic_plot)


# Filter data for TP2
tp2_data <- subset(Abundance_female_long, timepoint == "TP2")
print(tp2_data)
```
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity")
```


```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity")




plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity", position="fill", colour="black") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic()

# Display the plot at this point
print(plot)

plot <- plot + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line"),
          axis.text.x = element_text(angle = 65, hjust = 1, size = 60, color = "black"),
          axis.text.y = element_text(size = 60, color = "black"),
          axis.title.y = element_text(size = 70),
          plot.margin = margin(4, 4, 4, 4, "cm"))
```


```{r}
library(ggplot2)
library(grid)
library(RColorBrewer)

# Plot all
female_abund_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position="fill", stat="identity") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
    theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 60, color = "black"),
          axis.text.y = element_text(size = 60, color = "black"),
          axis.title.y = element_text(size = 70),
          plot.margin = margin(4, 4, 4, 4, "cm"))

##ggsave("output/Female_Gamete_Abundance_AST.pdf", female_abund_plot, width = 35, height = 40, units = "in")
##ggsave("output/Female_Gamete_Abundance_AST.png", female_abund_plot, width = 35, height = 40, units = "in")

```




female_abund_plot <- 

### Plot
```{r}
## Plot all 
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
geom_bar(position="fill", stat="identity", color = "black") +
scale_fill_manual(values = brewer.pal(n=5, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899" ), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4", "Stage 5")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm"))
#ggsave("output/Female_Gamete_Abundance_AST.pdf", female_abund_plot, width = 35, height = 40, units = "in")
#ggsave("output/Female_Gamete_Abundance_AST.png", female_abund_plot, width = 35, height = 40, units = "in")

## Plot by treatment 
# Reorder treatment variables 
Abundance_female_long$treatment <- factor(Abundance_female_long$treatment, levels = c("Ambient", "High", "Field"))

# Set up n size text - trying to add sample sizes under each of the bars, but it is proving difficult with facet_wrap
text_high <- text_grob("test")

# Plot
female_abund_trt_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=Timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
    facet_wrap(~Treatment) +
    #geom_bar(position="fill", stat="identity", color = "black") +
    #scale_fill_manual(values = brewer.pal(n=5, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0","#137899"), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4", "Stage 5")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_bw() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70), 
        axis.line = element_line(colour = "black", linewidth = 1)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm")) +
  theme(strip.text.x = element_text(size = 50, margin = margin(2, 0, 2, 0))) +
  #annotation_custom(text_high, xmin = 1, xmax=1,ymin=-0.07,ymax=-0.07) +
  coord_cartesian(clip = "off")
female_abund_trt_plot

ggsave("output/Female_Gamete_Abundance_Treatment.pdf", female_abund_trt_plot, width = 35, height = 36, units = "in")
ggsave("output/Female_Gamete_Abundance_Treatment.png", female_abund_trt_plot, width = 35, height = 36, units = "in")
```
