---
title: "AST_histo_abundance"
author: "Chloé Gilligan"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

This script will analyze gamete abundance in Astrangia poculata over time. The data is separated into 3 treatments (ambient temperature, high temperature, field). 


## Installing ggpubr package
```{r}
install.packages("ggpubr")
```

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("RColorBrewer")
library(ggpubr)
```

## Reading in data and metadata
```{r}
library(readr)
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
select(full_sample_id, timepoint, treatment, date, species, month, site,  sex, StageI, StageII, StageIII, StageIV, StageV, total_gametes)
##metadata <- read_csv("../AST_histology_metadata.csv")
```


## Filtering out the TP from sample ID's the begin with AST, keeping them on the field samples
```{r}
library(stringr)
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))
```

## Merging data to include treatment and site from the metadata to the histology abundance and size data
```{r}
##merged_data <- merge(data, metadata, by = "full_sample_id")
##write_csv(merged_data, "../data/merged_data.csv")
```

## Select females
```{r}
## changing merged_data to data once data was merged manually

Abundance_female <- data %>% 
  filter(sex == "female") 

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'
```


#Trying to see if I can do and if else statement and use sum of total gametes
```{r}
# Calculate the sum of Stages I-V for each row
#Abundance_female$sum_stages <- rowSums(Abundance_female[, c("Stage0", StageI", "StageII", "StageIII", "StageIV", "StageV")])

# Create a new column 'Stage' based on the sum
#Abundance_female$Stage <- ifelse(Abundance_female$sum_stages == 0, 0, "Other")

# Optionally, you can remove the 'sum_stages' column if you don't need it anymore
#Abundance_female$sum_stages <- NULL
```


### Calculate relative proportion of each stage 
```{r}
Abundance_female <- Abundance_female %>%
  mutate(#rel_prop_stage0 = (Stage0 / total_gametes) * 100,
         rel_prop_stage1 = (StageI / total_gametes) * 100,
         rel_prop_stage2 = (StageII / total_gametes) * 100, 
         rel_prop_stage3 = (StageIII / total_gametes) * 100, 
         rel_prop_stage4 = (StageIV / total_gametes) * 100,
         rel_prop_stage5 = (StageV / total_gametes) * 100)
```

### Duplicating data that is in ambient stage 0 goes into heat as well
```{r}
# Duplicate rows where treatment is "ambient"
ambient_data <- Abundance_female %>%
  filter(treatment == "Ambient")

# Change the treatment value to "heat"
ambient_data <- ambient_data %>%
  mutate(treatment = "Heat")

# Combine the original data with the duplicated data
Abundance_female <- bind_rows(Abundance_female, ambient_data)

```


### Troubleshooting to make if else statement to make Stage 0- not working
```{r}
# Find numeric columns with negative values
#numeric_cols <- sapply(Abundance_female, is.numeric)
#numeric_data <- Abundance_female[, numeric_cols]

# Replace negative values with 0 for numeric columns only
#Abundance_female[, numeric_cols] <- apply(numeric_data, 2, function(x) ifelse(x < 0, 0, x))

```


### Trying to make Stage 0- not working
```{r}
# Create a new column 'Stage' based on the sum of Stages I-V
#Abundance_female <- Abundance_female %>%
  #mutate(Stage = ifelse(rowSums(select(., starts_with("Stage"))) < 1, "Stage 0", "Other"))


# Calculate the total abundance for each row
#Abundance_female <- Abundance_female %>%
 # mutate(total_abundance = rowSums(select(., starts_with("rel_prop_stage"))))

# Calculate the relative abundance for each stage
#Abundance_female <- Abundance_female %>%
  #mutate_at(vars(starts_with("rel_prop_stage")), ~./total_abundance * 100)

# Calculate the remaining abundance for stage 0
#Abundance_female <- Abundance_female %>%
  #mutate(rel_prop_stage0 = 100 - rowSums(select(., starts_with("rel_prop_stage1"):starts_with("rel_prop_stage5"))))

# Replace negative values with 0
#Abundance_female[Abundance_female < 0] <- 0

```


### Changing data from wide to long
```{r}
Abundance_female_long <- Abundance_female %>%
  pivot_longer(cols = starts_with("rel_prop_stage"), 
               names_to = "stage", 
               values_to = "value") %>%
  mutate(stage = factor(stage, 
                        #levels = c("rel_prop_stage0", "rel_prop_stage1", "rel_prop_stage2", "rel_prop_stage3", "rel_prop_stage4", "rel_prop_stage5"), 
                        #labels = c("Stage0","Stage 1", "Stage 2", "Stage 3", "Stage 4", "Stage 5")))
levels = c("rel_prop_stage1", "rel_prop_stage2", "rel_prop_stage3", "rel_prop_stage4", "rel_prop_stage5"), 
                        labels = c("Stage 1", "Stage 2", "Stage 3", "Stage 4", "Stage 5")))
```

### Plotting Data
```{r}
## Take out Stage 0 to see proper dataset
# Using ggplot2 to create the plot
#ggplot(Abundance_female_long, aes(x = timepoint, y = value / 10, fill = stage)) +
  #geom_col() +
  #scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  #labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  #theme_minimal() +
  #scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    #labels = c( 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
#scale_fill_manual(values = c("#C4CACD", "#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    #labels = c('Stage 0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))

ggplot(Abundance_female_long, aes(x = timepoint, y = value / 10, fill = stage)) +
  stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
  #geom_col() +
  #scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
  scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c( 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
#scale_fill_manual(values = c("#C4CACD", "#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    #labels = c('Stage 0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))


```

### Filtering out TP3 and TP5
```{r}
# Separate data for different treatments
ambient_heat_data <- Abundance_female_long %>%
  filter(treatment %in% c("Ambient", "Heat")) %>%
  filter(!timepoint %in% c("TP3", "TP5"))

field_data <- Abundance_female_long %>%
  filter(treatment == "Field")

# Combine the datasets back together
combined_data <- bind_rows(ambient_heat_data, field_data)

```

### Plotting data by treatment
```{r}
# Now you can use ggplot2 to create the plot
ggplot(Abundance_female_long, aes(x = timepoint, y = value / 10, fill = stage)) +
  stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c( 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))+
  facet_wrap(~treatment)
#scale_fill_manual(values = c("#C4CACD", "#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                   # labels = c('Stage 0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
 # facet_wrap(~treatment)

```


### Trying to add Stage 0
```{r}
## Load packages 
library(tidyverse)
library(RColorBrewer)
library(ggpubr)

## Reading in data and metadata
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
  select(full_sample_id, timepoint, treatment, date, species, month, site, sex, StageI, StageII, StageIII, StageIV, StageV, total_gametes)

## Filtering out the TP from sample ID's that begin with AST, keeping them on the field samples
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))

## Select females
Abundance_female <- data %>% 
  filter(sex == "female") 

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'

### Calculate relative proportion of each stage 
Abundance_female <- Abundance_female %>%
  mutate(
    rel_prop_stage1 = (StageI / total_gametes) * 100,
    rel_prop_stage2 = (StageII / total_gametes) * 100, 
    rel_prop_stage3 = (StageIII / total_gametes) * 100, 
    rel_prop_stage4 = (StageIV / total_gametes) * 100,
    rel_prop_stage5 = (StageV / total_gametes) * 100
  )

# Calculate the sum of Stages I-V for each row
Abundance_female <- Abundance_female %>%
  rowwise() %>%
  mutate(sum_stages = sum(c(rel_prop_stage1, rel_prop_stage2, rel_prop_stage3, rel_prop_stage4, rel_prop_stage5), na.rm = TRUE)) %>%
  ungroup()

# Create Stage 0 based on the sum of other stages
Abundance_female <- Abundance_female %>%
  mutate(rel_prop_stage0 = ifelse(sum_stages < 100, 100 - sum_stages, 0))

# Changing data from wide to long
Abundance_female_long <- Abundance_female %>%
  pivot_longer(cols = starts_with("rel_prop_stage"), 
               names_to = "stage", 
               values_to = "value") %>%
  mutate(stage = factor(stage, 
                        levels = c("rel_prop_stage0", "rel_prop_stage1", "rel_prop_stage2", "rel_prop_stage3", "rel_prop_stage4", "rel_prop_stage5"), 
                        labels = c("Stage 0", "Stage 1", "Stage 2", "Stage 3", "Stage 4", "Stage 5")))

# Plotting Data
ggplot(Abundance_female_long, aes(x = timepoint, y = value, fill = stage)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
  scale_fill_manual(values = c("#C4CACD", "#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c('Stage 0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
```



### Troubleshooting- WORKS!!!
```{r}
# Load necessary libraries
library(tidyr)
library(dplyr)
library(ggplot2)

# Assuming Abundance_female is your original data frame with columns rel_prop_stage1 to rel_prop_stage5
# First, we need to reshape the data to long format

Abundance_female_long <- Abundance_female %>%
  pivot_longer(cols = starts_with("rel_prop_stage"), 
               names_to = "stage", 
               values_to = "value") %>%
  mutate(stage = factor(stage, 
                        levels = c("rel_prop_stage1", "rel_prop_stage2", "rel_prop_stage3", "rel_prop_stage4", "rel_prop_stage5"), 
                        labels = c("Stage 1", "Stage 2", "Stage 3", "Stage 4", "Stage 5")))

# Now you can use ggplot2 to create the plot
ggplot(Abundance_female_long, aes(x = timepoint, y = value / 10, fill = stage)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
  scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c( 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
#scale_fill_manual(values = c("#C4CACD", "#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                  #  labels = c('Stage 0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
```



### Trying to make sure stage 1 is included in TP0
```{r}
#Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>%
  #select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  #filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
  #na.omit()                    # Remove rows with NA in any other columns
```

### SUMMARIZE DATA
```{r}
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint) %>% 
  summarise(total_count = n())
```
#Trying to plot data
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(x = timepoint, y = value/10, fill = stage)) +
  geom_col() +  # Using geom_col() instead of geom_bar(stat = "identity")
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
  scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))

```

### Ifelse statement to turn anything that had zeros to stage 0 starting code from the top
```{r}
library(readr)
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
  select(full_sample_id, timepoint, treatment, date, species, month, site, sex, StageI, StageII, StageIII, StageIV, StageV, total_gametes)

library(stringr)
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))

#merged_data <- merge(data, metadata, by = "full_sample_id")
#write_csv(merged_data, "../data/merged_data.csv")

## changing merged_data to data once data was merged manually

Abundance_female <- data %>% 
  filter(sex == "female") 

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'

# Calculate the sum of Stages I-V for each row
Abundance_female$sum_stages <- rowSums(Abundance_female[, c("StageI", "StageII", "StageIII", "StageIV", "StageV")])

# Create a new column 'Stage' based on the sum
Abundance_female$Stage <- ifelse(Abundance_female$sum_stages == 0, 0, "Other")

# Optionally, you can remove the 'sum_stages' column if you don't need it anymore
Abundance_female$sum_stages <- NULL

Abundance_female <- Abundance_female %>%
  mutate(Stage0 = total_gametes - rowSums(Abundance_female[, c("StageI", "StageII", "StageIII", "StageIV", "StageV")]),
         rel_prop_stage0 = (Stage0 / total_gametes) * 100,
         rel_prop_stage1 = (StageI / total_gametes) * 100,
         rel_prop_stage2 = (StageII / total_gametes) * 100, 
         rel_prop_stage3 = (StageIII / total_gametes) * 100, 
         rel_prop_stage4 = (StageIV / total_gametes) * 100,
         rel_prop_stage5 = (StageV / total_gametes) * 100)

Abundance_female_long <- Abundance_female %>%
  pivot_longer(cols = rel_prop_stage0:rel_prop_stage5, 
               names_to = "stage", 
               values_to = "value") %>%
  select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  filter(!is.na(treatment)) %>%
  na.omit()

#Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5, key = "stage", value = "value")%>%
 # select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
 # filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
 # na.omit()                    # Remove rows with NA in any other columns


#agg_tbl_female <- Abundance_female %>%
  #group_by(treatment, timepoint) %>% 
 # summarise(total_count = n())

head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(x = timepoint, y = value/10, fill = stage)) +
  geom_col() +  # Using geom_col() instead of geom_bar(stat = "identity")
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal() +
  #scale_fill_manual(values = c("#C4CACD", #CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    #labels = c('Stage0', 'Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))



```




```{r}
### Trying to plot data
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(x = timepoint, y = value/100, fill = stage)) +
  geom_bar(stat="identity") +
  labs(x = NULL, y = "Relative Abundance", fill = "Stage") +
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
  #scale_fill_manual(values = c('#C4CACD',"#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      #labels = c('#Stage 0', 'Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
  
scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 1)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() +
scale_y_continuous(labels = scales::percent_format(scale = 1))


## Plot by treatment 
# Reorder treatment variables 
Abundance_female_long$treatment <- factor(Abundance_female_long$treatment, levels = c("Ambient", "High", "Field"))

# Set up n size text - trying to add sample sizes under each of the bars, but it is proving difficult with facet_wrap
text_high <- text_grob("test")
```



### Trying to plot data
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
  geom_bar(stat="identity") +
  labs(x = NULL, y = "Relative Abundance", fill = "Stage") +
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
  #scale_fill_manual(values = c('#C4CACD',"#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      #labels = c('#Stage 0', 'Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() +
scale_y_continuous(labels = scales::percent_format(scale = 1))


# Plot the data as a stacked bar graph
ggplot(Abundance_female_long, aes(x = timepoint, y = value/10, fill = stage)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()
```




```{r}
# Load necessary libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

# Filter out TP from sample IDs starting with AST
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))

# Select females and rename columns
Abundance_female <- data %>% 
  filter(sex == "female")

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'

# Calculate relative proportion of each stage
Abundance_female <- Abundance_female %>%
  mutate(
    rel_prop_stage1 = (StageI / total_gametes) * 100,
    rel_prop_stage2 = (StageII / total_gametes) * 100, 
    rel_prop_stage3 = (StageIII / total_gametes) * 100, 
    rel_prop_stage4 = (StageIV / total_gametes) * 100,
    rel_prop_stage5 = (StageV / total_gametes) * 100
  )

# Debugging: Check missing or zero values in rel_prop_stage1 for TP0
Abundance_female %>%
  filter(timepoint == "TP0") %>%
  select(full_sample_id, timepoint, StageI, total_gametes, rel_prop_stage1) %>%
  summarise(
    total_samples = n(),
    missing_values = sum(is.na(rel_prop_stage1)),
    zero_values = sum(rel_prop_stage1 == 0)
  )

# Debugging: Verify calculation of rel_prop_stage1
Abundance_female %>%
  filter(timepoint == "TP0") %>%
  mutate(
    calculated_rel_prop_stage1 = (StageI / total_gametes) * 100
  ) %>%
  select(full_sample_id, timepoint, StageI, total_gametes, rel_prop_stage1, calculated_rel_prop_stage1)

# Change data from wide to long
Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>%
  select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
  na.omit()                      # Remove rows with NA in any other columns

# Debugging: Verify filtering steps
Abundance_female_long %>%
  filter(timepoint == "TP0" & stage == "rel_prop_stage1")

# Summarize data
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint) %>% 
  summarise(total_count = n())

# Plot the data as a stacked bar graph
ggplot(Abundance_female_long, aes(x = timepoint, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()

#Trying to figure out where stage 1 went



# Plot the data as a stacked bar graph
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(x = timepoint, y = value/10, fill = stage)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()+
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5'))
```




```{r}
# Load necessary libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

# Filter out TP from sample IDs starting with AST
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))

# Select females and rename columns
Abundance_female <- data %>% 
  filter(sex == "female")

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'

# Calculate relative proportion of each stage
Abundance_female <- Abundance_female %>%
  mutate(
    rel_prop_stage1 = (StageI / total_gametes) * 100,
    rel_prop_stage2 = (StageII / total_gametes) * 100, 
    rel_prop_stage3 = (StageIII / total_gametes) * 100, 
    rel_prop_stage4 = (StageIV / total_gametes) * 100,
    rel_prop_stage5 = (StageV / total_gametes) * 100
  )

# Change data from wide to long
Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>%
  select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
  na.omit()                      # Remove rows with NA in any other columns

# Summarize data
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint) %>% 
  summarise(total_count = n())

# Plot the data as a stacked bar graph
ggplot(Abundance_female_long, aes(x = timepoint, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()


# Plot the data as a stacked bar graph
ggplot(Abundance_female_long, aes(x = timepoint, y = value/10, fill = stage)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()
```



```{r}
# Load necessary libraries
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

# Filter out TP from sample IDs starting with AST
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))

# Select females and rename columns
Abundance_female <- data %>% 
  filter(sex == "female")

names(Abundance_female)[2] <- 'timepoint'
names(Abundance_female)[3] <- 'treatment'
names(Abundance_female)[4] <- 'date'

# Calculate relative proportion of each stage
Abundance_female <- Abundance_female %>%
  mutate(
    total_count = StageI + StageII + StageIII + StageIV + StageV,
    rel_prop_stage1 = (StageI / total_gametes) * 100,
    rel_prop_stage2 = (StageII / total_gametes) * 100, 
    rel_prop_stage3 = (StageIII / total_gametes) * 100, 
    rel_prop_stage4 = (StageIV / total_gametes) * 100,
    rel_prop_stage5 = (StageV / total_gametes) * 100
  )

# Change data from wide to long
Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>%
  select(full_sample_id, treatment, timepoint, sex, stage, value) %>%
  filter(!is.na(treatment)) %>%  # Ensure treatment is not NA
  na.omit()                      # Remove rows with NA in any other columns

# Summarize data
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint) %>% 
  summarise(total_count = n())

# Plot the data
ggplot(Abundance_female_long, aes(x = timepoint, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Stages") +
  theme_minimal()


#Plotting by treatment
female_abund_trt_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=Timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
    facet_wrap(~Treatment) +
    #geom_bar(position="fill", stat="identity", color = "black") +
    #scale_fill_manual(values = brewer.pal(n=4, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0"), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_bw()


```

## Plotting Data
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
  geom_bar(stat="identity") +
  labs(x = NULL, y = "Relative Abundance", fill = "Stage") +
   scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))+
# scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() +
scale_y_continuous(labels = scales::percent_format(scale = 1))
#Figure out if relative porportion or not, or if stage 0 is needed?
theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm"))



```




```{r}
library(ggplot2)
library(RColorBrewer)

# Basic plot to check data
basic_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity", position="fill", colour="black") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() 

# Display the basic plot
print(basic_plot)


# Filter data for TP2
tp2_data <- subset(Abundance_female_long, timepoint == "TP2")
print(tp2_data)
```
```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity")
```


```{r}
head(Abundance_female_long)
str(Abundance_female_long)
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity")




plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    geom_bar(stat="identity", position="fill", colour="black") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic()

# Display the plot at this point
print(plot)

plot <- plot + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line"),
          axis.text.x = element_text(angle = 65, hjust = 1, size = 60, color = "black"),
          axis.text.y = element_text(size = 60, color = "black"),
          axis.title.y = element_text(size = 70),
          plot.margin = margin(4, 4, 4, 4, "cm"))
```


```{r}
library(ggplot2)
library(grid)
library(RColorBrewer)

# Plot all
female_abund_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position="fill", stat="identity") +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                      labels = c('Stage 1', 'Stage 2', 'Stage 3', 'Stage 4', 'Stage 5')) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
    theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 60, color = "black"),
          axis.text.y = element_text(size = 60, color = "black"),
          axis.title.y = element_text(size = 70),
          plot.margin = margin(4, 4, 4, 4, "cm"))

##ggsave("output/Female_Gamete_Abundance_AST.pdf", female_abund_plot, width = 35, height = 40, units = "in")
##ggsave("output/Female_Gamete_Abundance_AST.png", female_abund_plot, width = 35, height = 40, units = "in")

```




female_abund_plot <- 

### Plot
```{r}
## Plot all 
ggplot(Abundance_female_long, aes(fill=stage, y=value, x=timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
geom_bar(position="fill", stat="identity", color = "black") +
scale_fill_manual(values = brewer.pal(n=5, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899" ), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4", "Stage 5")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm"))
#ggsave("output/Female_Gamete_Abundance_AST.pdf", female_abund_plot, width = 35, height = 40, units = "in")
#ggsave("output/Female_Gamete_Abundance_AST.png", female_abund_plot, width = 35, height = 40, units = "in")

## Plot by treatment 
# Reorder treatment variables 
Abundance_female_long$treatment <- factor(Abundance_female_long$treatment, levels = c("Ambient", "High", "Field"))

# Set up n size text - trying to add sample sizes under each of the bars, but it is proving difficult with facet_wrap
text_high <- text_grob("test")

# Plot
female_abund_trt_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=Timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
    facet_wrap(~Treatment) +
    #geom_bar(position="fill", stat="identity", color = "black") +
    #scale_fill_manual(values = brewer.pal(n=5, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0","#137899"), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4", "Stage 5")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_bw() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70), 
        axis.line = element_line(colour = "black", linewidth = 1)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm")) +
  theme(strip.text.x = element_text(size = 50, margin = margin(2, 0, 2, 0))) +
  #annotation_custom(text_high, xmin = 1, xmax=1,ymin=-0.07,ymax=-0.07) +
  coord_cartesian(clip = "off")
female_abund_trt_plot

ggsave("output/Female_Gamete_Abundance_Treatment.pdf", female_abund_trt_plot, width = 35, height = 36, units = "in")
ggsave("output/Female_Gamete_Abundance_Treatment.png", female_abund_trt_plot, width = 35, height = 36, units = "in")
```
