---
title: "AST_histo_size"
author: "Chloé Gilligan"
date: "2024-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
```

#Reading in data
```{r}
Size <- read.csv("../data/astrangia_histology_size.csv", header = T)
```

#Calculating gemoetric mena and volume
```{r}
Size$geo.mean <- sqrt(Size$length*Size$width)

Size$geo.mean.vol <- pi*((Size$geo.mean/2)^2)
```

# Define standard error function
```{r}
std.error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}
```

# Separate full_sample_id and date
```{r}
Size <- Size %>%
  separate(full_sample_id, into = c("sample_id", "extra"), sep = "-", remove = FALSE) %>%
  separate(extra, into = c("id", "date"), sep = "-", remove = TRUE) #%>%
  #unite(full_sample_id, sample_id, id, sep = "-")
```


# Convert date to factors with desired labels
```{r}
Size <- Size %>%
  mutate(date = factor(date, 
                       levels = c("2/25/21","4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"), 
                       labels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")))
```

#Adding treatment
```{r}
Size <- Size %>%
  mutate(treatment = factor(treatment, levels = c("Field", "Ambient", "High")))
```


# Calculate means and standard errors
```{r}
size.means <- Size %>%
  group_by( date, month, sex, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop')
```

# Set factor levels for date and sex
```{r}
size.means$treatment <- factor(size.means$treatment, levels = c("field", "ambient", "high"))
size.means$date <- factor(size.means$date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
size.means$sex <- factor(size.means$sex, levels = c("male", "female"))
```

#Calculating volume means and stard errors
```{r}
vol.means <- Size %>%
  group_by( date, month, sex, treatment) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n())

vol.means[vol.means == "NaN"] <- NA
vol.means$month <- factor(vol.means$month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))

# vol.means$mm3 <- vol.means$gm / 10^9
```

```{r}
# Ensure the month variable is a factor with ordered levels
size.means <- size.means %>%
  mutate(month = factor(month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")))
```


# Create the plot
```{r}
# Plot all stages by Timepoint
size_plot <- ggplot(size.means, aes(x = timepoint, y = value, fill = stage)) +
  geom_point() +
  facet_wrap(~ stage)

# Plot stages by Timepoint and Treatment
stage_trt_plot <- ggplot(size.means, aes(x = timepoint, y = value, fill = stage)) +
  geom_point() +
  facet_grid(treatment ~ stage) +
  scale_fill_manual(
    values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
    labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V')) +
  theme_bw() +
  ylab("Microns (µm)") +
  xlab("Timepoint") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        panel.border = element_rect(colour = "black", size = 1),
        legend.position = "none")

print(size_plot)
print(stage_trt_plot)

# Save the plots
#ggsave("output/all_stages_plot.pdf", all_stages_plot, width = 10, height = 6, units = "in")
#ggsave("output/stage_trt_plot.pdf", stage_trt_plot, width = 10, height = 6, units = "in")
```



# Create the plot
Size.plot <- size.means %>%
  filter(!is.na(gm)) %>%
  ggplot(aes(x = month, y = gm, colour = treatment, label = treatment)) +
  geom_point(aes(), size = 2, position = position_dodge(width = 0.1)) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.1, position = position_dodge(width = 0.1)) +
  scale_color_manual(values = c("coral1", "blue")) +
  facet_wrap(~ treatment + sex, ncol = 1, scales = "free") +
  theme_bw() +
  ylab("Geometric Mean Diameter (µm)") +
  theme(strip.text.x = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top",
        axis.title.x = element_blank())

# Print the plot
Size.plot

```





```{r}
Size.plot <- size.means %>%
  filter(!is.na(gm)) %>%
  ggplot(aes(x = month, y = gm, colour = treatment, label = treatment)) +
  scale_x_discrete(limits = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")) +
  geom_point(aes(), size = 2, position = position_dodge(width = 0.1)) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.1, position = position_dodge(width = 0.1)) +
  scale_color_manual(values = c("coral1", "blue")) +
  facet_wrap(vars(sex, treatment), ncol = 1, scales = "free") +
  theme_bw() +
  ylab("Geometric Mean Diameter (µm)") +
  theme(strip.text.x = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(angle = 90),
        legend.position = "top",
        axis.title.x = element_blank())
Size.plot
```








