---
title: "AST_histo_size_field"
author: "Chloé Gilligan"
date: "2024-05-28"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
if (!require(gridExtra)) {
  install.packages("gridExtra")
  library(gridExtra)
}

```

#Reading in data
```{r}
Size <- read.csv("../data/astrangia_histology_size.csv", header = T, na.strings = c("", "NA")) %>% 
  select(full_sample_id, timepoint, date, treatment, sex, stage, length, width) %>% 
  filter(treatment %in% c("Field"))

print(head(Size))

```

#Calculating gemoetric mena and volume
```{r}
#Size$geo.mean <- sqrt(Size$length*Size$width)

#Size$geo.mean.vol <- pi*((Size$geo.mean/2)^2)
```

#Calculating gemoetric mena and volume
```{r}
Size_calc <- Size %>%
  mutate(
    geo.mean = sqrt(length * width),
    geo.mean.vol = pi * ((geo.mean / 2) ^ 2)
  )
print(head(Size_calc))

```

# Define standard error function
```{r}
std.error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}
```

#Fixing dates
```{r}
unique_dates <- unique(Size_calc$date)
print(unique_dates)
print(length(unique_dates))
```

# Calculate means and standard errors
```{r}
# Convert date column to factor with desired labels
size.means <- Size_calc %>%
  group_by(date, sex, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    date = factor(date, labels = c("4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"))
  )

# Ensure sex is treated as a factor
size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

print(size.means)

```


# Set factor levels for date and sex
```{r}
#size.means$treatment <- factor(size.means$treatment, levels = c("field", "ambient", "high"))
#size.means$date <- factor(size.means$date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

#size.means$date <- factor(size.means$date, levels = c("2/25/21","4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

```

```{r}
#size.means <- size.means %>%
  #mutate(
    #treatment = factor(treatment, levels = c("Field", "Ambient", "High")),
    #date = factor(date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")),
    #sex = factor(sex, levels = c("male", "female"))
  #)

#GETTING RID OF FOR ONLY A TEST
```

# Filter data for females and males
```{r}
size.means_female <- size.means %>% filter(sex == "female", treatment %in% c("Field"))
size.means_male <- size.means %>% filter(sex == "male", treatment %in% c("Field"))

size.means_female <- size.means %>% filter(sex == "female", treatment == "Field")
size.means_male <- size.means %>% filter(sex == "male", treatment == "Field")

print(size.means_female)
print(size.means_male)

```


#Calculating volume means and stard errors
```{r}
#vol.means <- Size %>%
  #group_by( date, month, sex, treatment) %>%
  #summarise(
    #gm = mean(geo.mean.vol, na.rm = TRUE), 
    #sem = std.error(geo.mean.vol),
    #n = n())

#vol.means[vol.means == "NaN"] <- NA
#vol.means$month <- factor(vol.means$month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))

# vol.means$mm3 <- vol.means$gm / 10^9
```


#Define a custom standard error function that handles NA values
```{r}
# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }
}

# Apply the custom function within summarise
```

#Volume
```{r}
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(gm = mean(geo.mean.vol, na.rm = TRUE), 
            sem = std.error(geo.mean.vol),
            n = n())

vol.means[vol.means == "NaN"] <- NA

#vol.means$mm3 <- vol.means$gm/10^9

vol.means$date <- factor(vol.means$date, levels = c("4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))

# Check the filtered data
print(size.means_filtered)

```


```{r}
# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }}

# Convert date column to Date format
Size_calc$date <- as.Date(Size_calc$date, format = "%m/%d/%y")

# Apply the custom function within summarise
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'
  )

# Ensure that any "NaN" values are replaced with NA
#vol.means[vol.means == "NaN"] <- NA

# Convert date column to factor with desired levels
vol.means$date <- factor(vol.means$date, levels = c("2021-04-11", "2021-04-29", "2021-06-04", "2021-06-26", "2021-07-24", "2021-09-02"))

```

#Plotting females

```{r}
plot_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = c("Field" = "blue")) +
  labs(title = "Geometric Mean Diameter for Females",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  )

print(plot_female)
```

#Plotting males
```{r}
plot_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = c("Field" = "blue")) +
  labs(title = "Geometric Mean Diameter for Males",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment"))
    
print(plot_male)
```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```




#Plotting Females
```{r}
plot_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_female$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Plotting Males
```{r}
plot_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Males",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_male$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```




