---
title: "AST_histo_size_field"
author: "Chloé Gilligan"
date: "2024-05-28"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
if (!require(gridExtra)) {
  install.packages("gridExtra")
  library(gridExtra)
}

```
#Cleaned up script
```{r}
# Load libraries
library(tidyverse)
library(gridExtra)
library(grid)

# Read and filter data
Size <- read.csv("../data/astrangia_histology_size.csv", header = TRUE, na.strings = c("", "NA")) %>%
  select(full_sample_id, timepoint, date, treatment, sex, stage, length, width) %>%
  filter(treatment == "Field")
```

```{r}
# Calculate geometric mean and volume
Size_calc <- Size %>%
  mutate(
    geo.mean = sqrt(length * width),
    geo.mean.vol = pi * ((geo.mean / 2) ^ 2)
  )
```

```{r}
# Standard error function
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) NA else sd(x, na.rm = TRUE) / sqrt(n)
}

# Calculate group means and SEs
size.means <- Size_calc %>%
  group_by(date, sex, stage, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE),
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    descriptive_date = factor(date,
      levels = c("4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"),
      labels = c("Early April", "Late April", "Early June", "Late June", "July", "September")
    ),
    sex = factor(sex, levels = c("male", "female")),
    stage = as.factor(stage)
  )
```

```{r}
# Split by sex
size.means_female <- filter(size.means, sex == "female")
size.means_male <- filter(size.means, sex == "male")

# Color palette
stage_colors <- c(
  "1" = "#CCEEF9",
  "2" = "#A6E1F4",
  "3" = "#59C7EB",
  "4" = "#00A9E0",
  "5" = "#137899"
)
```

```{r}
# Plot for females
plot_female_stage <- ggplot(size.means_female, aes(x = descriptive_date, y = gm, color = stage)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.2) +
  labs(
    title = "Female",
    x = "Date",
    y = "GMD (µm)",
    color = "Stage"
  ) +
  scale_color_manual(values = stage_colors) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Plot for males
plot_male_stage <- ggplot(size.means_male, aes(x = descriptive_date, y = gm, color = stage)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.2) +
  labs(
    title = "Male",
    x = "Date",
    y = "GMD (µm)",
    color = "Stage"
  ) +
  scale_color_manual(values = stage_colors) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Arrange with spacing
grid.arrange(
  plot_female_stage,
  plot_male_stage,
  ncol = 2,
  widths = c(1, 1),
  top = textGrob("Geometric Mean Diameter by Stage (Females vs. Males)", gp = gpar(fontsize = 16, fontface = "bold")),
  padding = unit(1, "lines")
)
```

#Stats work


```{r}
xtabs(~ sex + stage + date, data = Size_calc)
```

```{r}
Size_calc <- Size_calc %>%
  mutate(date_group = case_when(
    date %in% c("4/11/21", "4/29/21") ~ "April",
    date %in% c("6/4/21", "6/26/21") ~ "June",
    date == "7/24/21" ~ "July",
    date == "9/2/21" ~ "September"
  ),
  date_group = factor(date_group, levels = c("April", "June", "July", "September")))
```

```{r}
#Filtering dataset
Size_clean <- Size_calc %>%
  filter(!is.na(sex), stage != 0)

#Fit the model
lm_model <- lm(geo.mean ~ stage * sex + date, data = Size_clean)

# Type III ANOVA
library(car)
Anova(lm_model, type = "III")
```
```{r}
qqnorm(residuals(lm_model))
qqline(residuals(lm_model), col = "blue")
```


#Success!
```{r}
library(dplyr)



# Calculate mean size per sex and stage
stage_means <- Size_clean %>%
  group_by(sex, stage) %>%
  summarise(avg_size = mean(geo.mean, na.rm = TRUE), .groups = "drop") %>%
  arrange(sex, stage)

# Calculate percent change and absolute difference between consecutive stages
growth_by_stage <- stage_means %>%
  group_by(sex) %>%
  mutate(
    pct_change = (avg_size - lag(avg_size)) / lag(avg_size) * 100,
    abs_change = avg_size - lag(avg_size),
    from_stage = lag(stage),
    to_stage = stage
  ) %>%
  filter(!is.na(pct_change)) %>%
  arrange(sex, desc(pct_change))  # sort by largest growth

growth_by_stage
```
#Success!
```{r}
# Load emmeans package for post-hoc comparisons
library(emmeans)

# Use the model you already fit
lm_model <- lm(geo.mean ~ stage * sex + date, data = Size_clean)

# Pairwise comparisons between stages, separately by sex
emmeans_stage <- emmeans(lm_model, ~ stage | sex)

# Compare consecutive stages (e.g., 1 vs 2, 2 vs 3, etc.)
pairs(emmeans_stage, adjust = "tukey")
```






#Stuff that did not work stats

```{r}
library(car)

# Ensure factors are treated correctly
Size_calc <- Size_calc %>%
  mutate(
    stage = as.factor(stage),
    sex = as.factor(sex),
    date = as.factor(date)
  )

# Linear model with interaction


lm_model1 <- lm(geo.mean ~ stage + sex + date + stage:sex + stage:date, data = Size_calc)
Anova(lm_model1, type = "III")

lm_model2 <- lm(geo.mean ~ stage * sex * date, data = Size_calc)
Anova(lm_model2, type = "III")
```



```{r}
size_stage_means <- Size_calc %>%
  group_by(sex, date, stage) %>%
  summarise(avg_size = mean(geo.mean, na.rm = TRUE), .groups = "drop") %>%
  arrange(sex, date, stage)

# Compute percent change between consecutive stages
size_pct_change <- size_stage_means %>%
  group_by(sex, date) %>%
  mutate(
    pct_change = (avg_size - lag(avg_size)) / lag(avg_size) * 100,
    stage_diff = stage - lag(stage)
  ) %>%
  filter(!is.na(pct_change) & stage_diff == 1)
```

```{r}
lm_model <- lm(pct_change ~ sex * date, data = size_pct_change)
summary(lm_model)
```

```{r}
library(car)
Anova(lm_model, type = "III")
```



```{r}
# Ensure stage is numeric for calculation
Size_calc <- Size_calc %>% mutate(stage = as.numeric(stage))

size_pct_change <- Size_calc %>%
  group_by(sex, date) %>%
  arrange(stage, .by_group = TRUE) %>%
  summarise(
    avg_size = mean(geo.mean, na.rm = TRUE),
    stage = stage,
    .groups = "drop_last"
  ) %>%
  mutate(
    pct_change = (avg_size - lag(avg_size)) / lag(avg_size) * 100,
    stage_diff = stage - lag(stage)
  ) %>%
  filter(!is.na(pct_change) & stage_diff == 1)

lm_model <- lm(pct_change ~ sex * date, data = size_pct_change)
summary(lm_model)

lm_model <- lm(pct_change ~ sex * stage_diff * date, data = size_pct_change)

library(car)
Anova(lm_model, type = "III")


```


```{r}
# Calculate the average size (geo.mean) for each stage, date, and sex
size_means <- Size_calc %>%
  group_by(sex, stage, date) %>%
  summarise(avg_size = mean(geo.mean, na.rm = TRUE), .groups = "drop")
```

```{r}
# Step 1: Calculate percent change and create 'stage_diff'
size_pct_change <- size_means %>%
  arrange(sex, date, stage) %>%
  group_by(sex, date) %>%
  mutate(
    pct_change = (avg_size - lag(avg_size)) / lag(avg_size) * 100,
    stage_diff = stage - lag(stage)  # This identifies the consecutive stage transitions
  ) %>%
  filter(!is.na(pct_change) & stage_diff == 1)  # Only include consecutive stage transitions

# Step 2: Run Type III ANOVA (with 'stage_diff')
anova_model_simple <- aov(pct_change ~ sex * stage_diff * date, data = size_pct_change)

# Step 3: Perform Type III ANOVA using the 'Anova' function from the 'car' package
library(car)  # Make sure 'car' package is loaded
anova_model_simple <- aov(pct_change ~ sex * stage_diff + date, data = size_pct_change)
Anova(anova_model_simple, type = "III")
```









```{r}
size.pctchange <- size.means %>%
  arrange(sex, date, treatment, as.numeric(as.character(stage))) %>%  # Ensure stage is numeric for ordering
  group_by(sex, date, treatment) %>%
  mutate(
    pct_change = (gm - lag(gm)) / lag(gm) * 100
  ) %>%
  ungroup()
```

```{r}
lm_simple <- lm(pct_change ~ sex + stage, data = size.pctchange)
anova(lm_simple)
summary(lm_simple)
```
```{r}
install.packages("car")
library(car)
```

```{r}
model_type3 <- lm(pct_change ~ sex + stage + date, data = size.pctchange)

Anova(model_type3, type = "III")
```


```{r}
lm(pct_change ~ sex + stage + date, data = size.pctchange)
model_time <- lm(pct_change ~ sex + stage + date, data = size.pctchange)
summary(model_time)
anova(model_time)
```


##Below here was the creation/ draft

#Reading in data
```{r}
Size <- read.csv("../data/astrangia_histology_size.csv", header = T, na.strings = c("", "NA")) %>% 
  select(full_sample_id, timepoint, date, treatment, sex, stage, length, width) %>% 
  filter(treatment %in% c("Field"))

print(head(Size))

```

#Calculating gemoetric mean and volume
```{r}
Size_calc <- Size %>%
  mutate(
    geo.mean = sqrt(length * width),
    geo.mean.vol = pi * ((geo.mean / 2) ^ 2)
  )
print(head(Size_calc))

```

# Define standard error function
```{r}
std.error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}
```

#Fixing dates
```{r}
unique_dates <- unique(Size_calc$date)
print(unique_dates)
print(length(unique_dates))
```

#Trying to plot with dates
```{r}
size.means <- Size_calc %>%
  group_by(date, sex, stage, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    descriptive_date = factor(date, 
                              levels = c("4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"),
                              labels = c("Early April", "Late April", "Early June", "Late June", "July", "September"))
  )

size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

print(size.means)

```


# Calculate means and standard errors
```{r}
# Convert date column to factor with desired labels
size.means <- Size_calc %>%
  group_by(date, sex, stage, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    date = factor(date, labels = c("4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"))
  )

# Ensure sex is treated as a factor
size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

print(size.means)

```


# Set factor levels for date and sex
```{r}
#size.means$treatment <- factor(size.means$treatment, levels = c("field", "ambient", "high"))
#size.means$date <- factor(size.means$date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

#size.means$date <- factor(size.means$date, levels = c("2/25/21","4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

```

# Filter data for females and males
```{r}
size.means_female <- size.means %>% filter(sex == "female", treatment %in% c("Field"))
size.means_male <- size.means %>% filter(sex == "male", treatment %in% c("Field"))

size.means_female <- size.means %>% filter(sex == "female", treatment == "Field")
size.means_male <- size.means %>% filter(sex == "male", treatment == "Field")

# Ensure sex is treated as a factor
size.means_female$sex <- factor(size.means_female$sex, levels = c("male", "female"))
size.means_male$sex <- factor(size.means_male$sex, levels = c("male", "female"))

# Ensure stage is treated as a factor
size.means_female$stage <- factor(size.means_female$stage)
size.means_male$stage <- factor(size.means_male$stage)


print(size.means_female)
print(size.means_male)

```


#Calculating volume means and stard errors
```{r}
#vol.means <- Size %>%
  #group_by( date, month, sex, treatment) %>%
  #summarise(
    #gm = mean(geo.mean.vol, na.rm = TRUE), 
    #sem = std.error(geo.mean.vol),
    #n = n())

#vol.means[vol.means == "NaN"] <- NA
#vol.means$month <- factor(vol.means$month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))

# vol.means$mm3 <- vol.means$gm / 10^9
```


#Define a custom standard error function that handles NA values 
```{r}
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }}

```

#Volume calculations
```{r}
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(gm = mean(geo.mean.vol, na.rm = TRUE), 
            sem = std.error(geo.mean.vol),
            n = n())

# Apply the custom function within summarise
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'
  )

vol.means[vol.means == "NaN"] <- NA

#vol.means$mm3 <- vol.means$gm/10^9

vol.means$date <- factor(vol.means$date, levels = c("4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))

# Convert date column to factor with desired levels
vol.means$date <- factor(vol.means$date, levels = c("2021-04-11", "2021-04-29", "2021-06-04", "2021-06-26", "2021-07-24", "2021-09-02"))

# Check the filtered data
print(size.means_filtered)

```

#Plotting test

```{r}
# Your provided ggplot code with modifications
plot_female_stage <- ggplot(size.means_female, aes(x = descriptive_date, y = gm, color = stage)) +
  geom_point(size = 3, shape = 16) + # shape = 16 ensures circles
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.2) +
  labs(
    title = "GMD for Females by Stage",
    x = "Date",
    y = "Geometric Mean Diameter(µm)",
    color = "Stage"
  ) +
  scale_color_manual(values = c(
    "1" = "#CCEEF9",
    "2" = "#A6E1F4",
    "3" = "#59C7EB",
    "4" = "#00A9E0",
    "5" = "#137899"
  )) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(plot_female_stage)

```
#Plotting males
```{r}
# Your provided ggplot code with modifications for the male plot
plot_male_stage <- ggplot(size.means_male, aes(x = descriptive_date, y = gm, color = stage)) +
  geom_point(size = 3, shape = 16) + # shape = 16 ensures circles
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.2) +
  labs(
    title = "GMD for Males by Stage",
    x = "Date",
    y = "Geometric Mean Diameter (µm)",
    color = "Stage"
  ) +
  scale_color_manual(values = c(
    "1" = "#CCEEF9",
    "2" = "#A6E1F4",
    "3" = "#59C7EB",
    "4" = "#00A9E0",
    "5" = "#137899"
  )) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(plot_male_stage)
```

```{r}
grid.arrange(plot_female_stage, plot_male_stage, ncol = 2)
```






```{r}
plot_female_stage <- ggplot(size.means_female, aes(x = date, y = gm, shape = stage, color = stage)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.2) +
  labs(title = "Geometric Mean Diameter for Females by Stage",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Stage",
       color = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_female_stage)

```





```{r}
#Plotting test
plot_stage_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
  scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
  labs(title = "Geometric Mean Diameter by Stage - Female",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Stage",
       linetype = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Stage"),
    linetype = guide_legend(title = "Stage")
  )

plot_stage_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
 scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
  labs(title = "Geometric Mean Diameter by Stage - Male",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Stage",
       linetype = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Stage"),
    linetype = guide_legend(title = "Stage")
  )

# Print both plots
print(plot_stage_female)
print(plot_stage_male)


```
#Plotting males and females together
```{r}
# Combine the two plots
combined_plot <- plot_stage_female / plot_stage_male

# Print the combined plot
print(combined_plot)
```



#Plotting test
```{r}
# Plotting test
plot_stage_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
  scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
  labs(title = "Geometric Mean Diameter by Stage - Female",
       x = "Date",
       y = "Geometric Mean Diameter",
       color = "Stage",  # Removed shape legend title
       linetype = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Stage"),
    linetype = guide_legend(title = "Stage")
  )

plot_stage_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
  scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
  labs(title = "Geometric Mean Diameter by Stage - Male",
       x = "Date",
       y = "Geometric Mean Diameter",
       color = "Stage",  # Removed shape legend title
       linetype = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Stage"),
    linetype = guide_legend(title = "Stage")
  )

# Print both plots
print(plot_stage_female)
print(plot_stage_male)

```





#Plotting 
```{r}
# Check if the number of unique dates is greater than a threshold to determine whether to plot all data or create box and whisker plots
if (length(unique(size.means_female$date)) > threshold) {
  plot_stage_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
    geom_point(size = 3) +
    geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
    scale_shape_manual(values = c("Field" = 16)) +
    scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
    scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
    labs(title = "Geometric Mean Diameter by Stage - Female",
         x = "Date",
         y = "Geometric Mean Diameter",
         shape = "Treatment",
         color = "Stage",
         linetype = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(
      shape = guide_legend(title = "Treatment"),
      color = guide_legend(title = "Stage"),
      linetype = guide_legend(title = "Stage")
    )
} else {
  plot_stage_female <- ggplot(size.means_female, aes(x = date, y = gm, fill = as.factor(stage))) +
    geom_boxplot() +
    labs(title = "Box and Whisker Plot of Geometric Mean Diameter by Stage - Female",
         x = "Date",
         y = "Geometric Mean Diameter",
         fill = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Similarly, create a plot for males
if (length(unique(size.means_male$date)) > threshold) {
  plot_stage_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
    geom_point(size = 3) +
    geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
    scale_shape_manual(values = c("Field" = 16)) +
    scale_color_manual(values = rev(c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9"))) + # Reverse color order for stages
    scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
    labs(title = "Geometric Mean Diameter by Stage - Male",
         x = "Date",
         y = "Geometric Mean Diameter",
         shape = "Treatment",
         color = "Stage",
         linetype = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(
      shape = guide_legend(title = "Treatment"),
      color = guide_legend(title = "Stage"),
      linetype = guide_legend(title = "Stage")
    )
} else {
  plot_stage_male <- ggplot(size.means_male, aes(x = date, y = gm, fill = as.factor(stage))) +
    geom_boxplot() +
    labs(title = "Box and Whisker Plot of Geometric Mean Diameter by Stage - Male",
         x = "Date",
         y = "Geometric Mean Diameter",
         fill = "Stage") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Print both plots
print(plot_stage_female)
print(plot_stage_male)

```





#Plotting with stages
```{r}
plot_stage <- ggplot(size.means_stage, aes(x = date, y = gm, shape = treatment, color = as.factor(stage), linetype = stage)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(stage, treatment)), size = 1) + # Add lines connecting points for each stage
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = c("#137899", "#00A9E0", "#59C7EB", "#A6E1F4", "#CCEEF9")) + # Adjust colors for stages
  scale_linetype_manual(values = c("Stage1" = "solid", "Stage2" = "dashed", "Stage3" = "dotted")) + # Adjust linetypes as needed
  labs(title = "Geometric Mean Diameter by Stage",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Stage",
       linetype = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Stage"),
    linetype = guide_legend(title = "Stage")
  )

print(plot_stage)

```






#Plotting females

```{r}
plot_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = c("Field" = "blue")) +
  labs(title = "Geometric Mean Diameter for Females",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  )

print(plot_female)
```

#Plotting males
```{r}
plot_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Field" = 16)) +
  scale_color_manual(values = c("Field" = "blue")) +
  labs(title = "Geometric Mean Diameter for Males",
       x = "Date",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment"))
    
print(plot_male)
```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```




#Plotting Females
```{r}
plot_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_female$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Plotting Males
```{r}
plot_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Males",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_male$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```




