---
title: "AST_histo_abundance"
author: "Chloé Gilligan"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

This script will analyze gamete abundance in Astrangia poculata over time. The data is separated into 3 treatments (ambient temperature, high temperature, field). 


## Installing ggpubr package
```{r}
install.packages("ggpubr")
```

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("RColorBrewer")
library(ggpubr)
```

## Reading in data and metadata
```{r}
library(readr)
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
select(full_sample_id, timepoint, date, species, month, site,  sex, StageI, StageII, StageIII, StageIV, StageV, total_gametes)
metadata <- read_csv("../AST_histology_metadata.csv")
```


## Filtering out the TP from sample ID's the begin with AST, keeping them on the field samples
```{r}
library(stringr)
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))
```

## Merging data to include treatment and site from the metadata to the histology abundance and size data
```{r}
merged_data <- merge(data, metadata, by = "full_sample_id")
write_csv(merged_data, "../data/merged_data.csv")
```

## Select females
```{r}
Abundance_female <- merged_data %>% 
  filter(sex == "female") 
```

### Calculate relative proportion of each stage 
```{r}
Abundance_female <- Abundance_female %>%
  mutate(rel_prop_stage1 = (StageI / total_gametes) * 100,
         rel_prop_stage2 = (StageII / total_gametes) * 100, 
         rel_prop_stage3 = (StageIII / total_gametes) * 100, 
         rel_prop_stage4 = (StageIV / total_gametes) * 100,
         rel_prop_stage5 = (StageV / total_gametes) * 100)
```


### Change data from wide to long 
```{r}
Abundance_female_long <- gather(Abundance_female, stage, value, rel_prop_stage1:rel_prop_stage5) %>% 
  select(full_sample_id, treatment, timepoint.x, sex, stage, value) %>% 
  na.omit()

```

### SUMMARIZE DATA
```{r}
agg_tbl_female <- Abundance_female %>%
  group_by(treatment, timepoint.x) %>% 
  summarise(total_count = n())
```

### Plot
```{r}
## Plot all 
female_abund_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=Timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
    #geom_bar(position="fill", stat="identity", color = "black") +
    #scale_fill_manual(values = brewer.pal(n=4, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0"), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_classic() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm"))
ggsave("output/Female_Gamete_Abundance_AST.pdf", female_abund_plot, width = 35, height = 40, units = "in")
ggsave("output/Female_Gamete_Abundance_AST.png", female_abund_plot, width = 35, height = 40, units = "in")

## Plot by treatment 
# Reorder treatment variables 
Abundance_female_long$Treatment <- factor(Abundance_female_long$Treatment, levels = c("Ambient", "High", "Field"))

# Set up n size text - trying to add sample sizes under each of the bars, but it is proving difficult with facet_wrap
text_high <- text_grob("test")

# Plot
female_abund_trt_plot <- ggplot(Abundance_female_long, aes(fill=stage, y=value, x=Timepoint)) + 
    stat_summary(fun.y=sum, geom="bar", colour="black", position = "fill", stat = "identity") +
    facet_wrap(~Treatment) +
    #geom_bar(position="fill", stat="identity", color = "black") +
    #scale_fill_manual(values = brewer.pal(n=4, name = "Blues")) +
    scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0"), labels=c('Stage 1', 'Stage 2', "Stage 3", "Stage 4")) +
    scale_y_continuous(expand = c(0, 0)) +
    ylab("Relative Proportion") +
    xlab("") +
    theme_bw() + 
    theme(legend.title = element_text(size = 0),
          legend.text = element_text(size = 50),
          legend.position = "right",
          legend.key.size = unit(3, "line")) +
          #legend.direction = "horizontal") +
    theme(axis.text.x = element_text(angle = 65,
                                   hjust = 1,
                                   size = 60,
                                   color = "black")) +
   theme(axis.text.y = element_text(size = 60,
                                   color = "black"),
        axis.title.y = element_text(size = 70), 
        axis.line = element_line(colour = "black", linewidth = 1)) +
  theme(plot.margin = margin(4, 4, 4, 4, "cm")) +
  theme(strip.text.x = element_text(size = 50, margin = margin(2, 0, 2, 0))) +
  #annotation_custom(text_high, xmin = 1, xmax=1,ymin=-0.07,ymax=-0.07) +
  coord_cartesian(clip = "off")
female_abund_trt_plot

ggsave("output/Female_Gamete_Abundance_Treatment.pdf", female_abund_trt_plot, width = 35, height = 36, units = "in")
ggsave("output/Female_Gamete_Abundance_Treatment.png", female_abund_trt_plot, width = 35, height = 36, units = "in")
```
