---
title: "AST_histo_size"
author: "Chloé Gilligan"
date: "2024-05-21"
output: html_document
---


```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
if (!require(gridExtra)) {
  install.packages("gridExtra")
  library(gridExtra)
}

```

#Reading in data
```{r}
Size <- read.csv("../data/astrangia_histology_size.csv", header = T, na.strings = c("", "NA")) %>% 
  select(full_sample_id, timepoint, date, treatment, sex, stage, length, width) %>% 
  filter(treatment %in% c("Ambient", "Heat"))
```


#Calculating gemoetric mena and volume
```{r}
Size_calc <- Size %>%
  mutate(
    geo.mean = sqrt(length * width),
    geo.mean.vol = pi * ((geo.mean / 2) ^ 2)
  )
```


# Define standard error function
```{r}
std.error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}
```

#Fixing dates
```{r}
unique_dates <- unique(Size_calc$date)
print(unique_dates)
print(length(unique_dates))
```



# Calculate means and standard errors
```{r}
# Convert date column to factor with desired labels
size.means <- Size_calc %>%
  group_by(date, sex, treatment) %>%
  summarise(
    gm = mean(geo.mean, na.rm = TRUE), 
    sem = std.error(geo.mean),
    n = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    date = factor(date, labels = c("2/25/21", "4/2/21", "6/2/21", "7/23/21", "8/25/21")))


# Ensure sex is treated as a factor
size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

```



# Set factor levels for date and sex
```{r}
#size.means$treatment <- factor(size.means$treatment, levels = c("field", "ambient", "high"))
#size.means$date <- factor(size.means$date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

#size.means$date <- factor(size.means$date, levels = c("2/25/21","4/11/21", "4/29/21","6/4/21", "6/26/21","7/24/21", "9/2/21"))
#size.means$sex <- factor(size.means$sex, levels = c("male", "female"))

```

```{r}
#size.means <- size.means %>%
  #mutate(
    #treatment = factor(treatment, levels = c("Field", "Ambient", "High")),
    #date = factor(date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")),
    #sex = factor(sex, levels = c("male", "female"))
  #)

#GETTING RID OF FOR ONLY A TEST
```

# Filter data for females
```{r}
#size.means_female <- size.means %>% filter(sex == "female")

#size.means_female <- size.means %>% filter(sex == "female", treatment %in% c("Ambient", "Heat"))
#size.means_male <- size.means %>% filter(sex == "male", treatment %in% c("Ambient", "Heat"))

size.means_female <- size.means %>% filter(sex == "female", treatment %in% c("Ambient", "Heat"))
size.means_male <- size.means %>% filter(sex == "male", treatment %in% c("Ambient", "Heat"))

```


#Calculating volume means and stard errors
```{r}
#vol.means <- Size %>%
  #group_by( date, month, sex, treatment) %>%
  #summarise(
    #gm = mean(geo.mean.vol, na.rm = TRUE), 
    #sem = std.error(geo.mean.vol),
    #n = n())

#vol.means[vol.means == "NaN"] <- NA
#vol.means$month <- factor(vol.means$month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))

# vol.means$mm3 <- vol.means$gm / 10^9
```


#Volume
```{r}
#vol.means <- Size_calc %>%
 # group_by(date, sex, treatment) %>%
 # summarise(
  #  gm = mean(geo.mean.vol, na.rm = TRUE), 
   # sem = std.error(geo.mean.vol),
   # n = n(),
   # .groups = 'drop'
#  )

#vol.means <- vol.means %>%
#  mutate(
   # date = factor(date, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September"))
 # )

#vol.means <- Size_calc %>%
  #group_by(treatment, date, sex) %>%
  #summarise(gm = mean(geo.mean.vol, na.rm=TRUE), 
            #sem = std.error(geo.mean.vol, na.rm=TRUE),
            #n=n())


# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }
}

# Apply the custom function within summarise
```

#Volume for experimental script- Not working anymore
```{r}
#vol.means <- Size_calc %>%
 # group_by(treatment, date, sex) %>%
 # summarise(gm = mean(geo.mean.vol, na.rm = TRUE), 
          #  sem = std.error(geo.mean.vol),
          #  n = n())

#vol.means[vol.means == "NaN"] <- NA

#vol.means$mm3 <- vol.means$gm/10^9

#vol.means$date <- factor(vol.means$date, levels = c("2/25/21", "4/2/21", "6/2/21", "7/23/21", "8/25/21"))

# Check the filtered data
#print(size.means_filtered)

```
#Volume for experimental script
```{r}
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'  # This ensures the grouped output is not grouped further
  ) %>%
  mutate(
    gm = ifelse(is.nan(gm), NA, gm),
    sem = ifelse(is.nan(sem), NA, sem)
  )

# Ensure date column is treated as a factor with the correct order
vol.means$date <- factor(vol.means$date, levels = c("2/25/21", "4/2/21", "6/2/21", "7/23/21", "8/25/21"))

# Print the resulting dataframe
print(vol.means)

```


```{r}
# Define a custom standard error function that handles NA values
std.error <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) {
    return(NA)
  } else {
    return(sd(x, na.rm = TRUE) / sqrt(n))
  }}

# Convert date column to Date format
Size_calc$date <- as.Date(Size_calc$date, format = "%m/%d/%y")

# Apply the custom function within summarise
vol.means <- Size_calc %>%
  group_by(treatment, date, sex) %>%
  summarise(
    gm = mean(geo.mean.vol, na.rm = TRUE), 
    sem = std.error(geo.mean.vol),
    n = n(),
    .groups = 'drop'
  )

# Ensure that any "NaN" values are replaced with NA
#vol.means[vol.means == "NaN"] <- NA

# Convert date column to factor with desired levels
vol.means$date <- factor(vol.means$date, levels = c("2021-02-25", "2021-04-02", "2021-06-02", "2021-07-23", "2021-08-25"))

```

#Plotting Males and Females
```{r}
plot_female <- ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_female$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Plotting Males
```{r}
plot_male <- ggplot(size.means_male, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Males",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  scale_y_continuous(breaks = seq(0, max(size.means_male$gm, na.rm = TRUE), by = 10))  # Adjust the 'by' argument to change spacing

```

#Putting plots side by side
```{r}
grid.arrange(plot_female, plot_male, ncol = 2)
```



#Plotting
```{r}
# Filter data for females and for only ambient and heat treatments
size.means_filtered <- size.means %>% 
  filter(sex == "female", treatment %in% c("Ambient", "Heat"))

# Generate the plot
ggplot(size.means_filtered, aes(x = date, y = gm, shape = treatment, group = treatment)) +
  geom_point(aes(shape = treatment), size = 3) +
  geom_line(aes(linetype = treatment), size = 1) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
c


```
```{r}
# Filter data for females and for only ambient and heat treatments
size.means_filtered <- size.means %>% 
  filter(sex == "female", treatment %in% c("Ambient", "Heat"))

# Generate the plot
ggplot(size.means_filtered, aes(x = date, y = gm, shape = treatment, color = treatment)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_color_manual(values = c("Ambient" = "blue", "Heat" = "red")) +  # blue for Ambient, red for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       color = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Combine legend for shapes and colors
  guides(
    shape = guide_legend(title = "Treatment"),
    color = guide_legend(title = "Treatment")
  ) +
  # Adjust y-axis ticks and labels
  scale_y_continuous(breaks = seq(0, max(size.means_filtered$gm), by = 10))  # Adjust the 'by' argument to change spacing

```




```{r}
# Filter data for females and for only ambient and heat treatments
size.means_filtered <- size.means %>% 
  filter(sex == "female", treatment %in% c("Ambient", "Heat"))

# Generate the plot
ggplot(size.means_filtered, aes(x = date, y = gm, shape = treatment, linetype = treatment)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(treatment, date)), size = 1) +  # Add line connecting shapes
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_linetype_manual(values = c("Ambient" = "solid", "Heat" = "dashed")) +
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       linetype = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Combine legend for shapes and lines
  guides(
    shape = guide_legend(title = "Treatment", override.aes = list(linetype = c("solid", "dashed")))
  )

```

```{r}
# Filter data for females and for only ambient and heat treatments
size.means_filtered <- size.means %>% 
  filter(sex == "female", treatment %in% c("Ambient", "Heat"))

# Generate the plot
ggplot(size.means_filtered, aes(x = date, y = gm, shape = treatment, linetype = treatment)) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(treatment, date)), size = 1) +  # Add line connecting shapes
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  scale_linetype_manual(values = c("Ambient" = "solid", "Heat" = "dashed")) +
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment",
       linetype = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Combine legend for shapes and lines
  guides(
    shape = guide_legend(title = "Treatment", override.aes = list(linetype = c("solid", "dashed")))
  )

```



```{r}
# Filter data for females and for only ambient and heat treatments
size.means_filtered <- size.means %>% 
  filter(sex == "female", treatment %in% c("Ambient", "Heat"))

# Generate the plot
ggplot(size.means_filtered, aes(x = date, y = gm, shape = treatment, group = treatment)) +
  geom_point(aes(shape = treatment), size = 3) +
  geom_line(aes(linetype = treatment), size = 1) +
  scale_shape_manual(values = c("Ambient" = 16, "Heat" = 17)) +  # circles for Ambient, triangles for Heat
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



```{r}
# Generate the plot
ggplot(size.means_female, aes(x = date, y = gm, shape = treatment, group = treatment)) +
  geom_point(aes(shape = treatment), size = 3) +
  geom_line(aes(linetype = treatment), size = 1) +
  scale_shape_manual(values = c("Ambient" = 16, "High" = 17)) +  # circles for Ambient, triangles for High
  labs(title = "Geometric Mean Diameter for Females",
       x = "Months",
       y = "Geometric Mean Diameter",
       shape = "Treatment") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




```{r}
# Ensure the month variable is a factor with ordered levels
#size.means <- size.means %>%
  #mutate(month = factor(month, levels = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")))

# Is month right here or should I be doing date? and then turning date into month after?
```


# Create the plot





```{r}
# Plot all stages by Timepoint
size_plot <- ggplot(size.means, aes(x = timepoint, y = value, fill = stage)) +
  geom_point() +
  facet_wrap(~ treatment)
# replaced treatment for stage

# Plot stages by Timepoint and Treatment
stage_trt_plot <- ggplot(size.means, aes(x = timepoint, y = value, fill = stage)) +
  geom_point() +
  facet_grid(treatment ~ stage) +
  scale_fill_manual(
    values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
    labels = c('Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V')) +
  theme_bw() +
  ylab("Microns (µm)") +
  xlab("Timepoint") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 10),
        panel.border = element_rect(colour = "black", size = 1),
        legend.position = "none")

print(size_plot)
print(stage_trt_plot)

# Save the plots
#ggsave("output/all_stages_plot.pdf", all_stages_plot, width = 10, height = 6, units = "in")
#ggsave("output/stage_trt_plot.pdf", stage_trt_plot, width = 10, height = 6, units = "in")
```



# Create the plot
Size.plot <- size.means %>%
  filter(!is.na(gm)) %>%
  ggplot(aes(x = month, y = gm, colour = treatment, label = treatment)) +
  geom_point(aes(), size = 2, position = position_dodge(width = 0.1)) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.1, position = position_dodge(width = 0.1)) +
  scale_color_manual(values = c("coral1", "blue")) +
  facet_wrap(~ treatment + sex, ncol = 1, scales = "free") +
  theme_bw() +
  ylab("Geometric Mean Diameter (µm)") +
  theme(strip.text.x = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top",
        axis.title.x = element_blank())

# Print the plot
Size.plot

```


```{r}
Size.plot <- size.means %>%
  filter(!is.na(gm)) %>%
  ggplot(aes(x = month, y = gm, colour = treatment, label = treatment)) +
  scale_x_discrete(limits = c("February", "Early April", "Late April", "Early June", "Late June", "July", "September")) +
  geom_point(aes(), size = 2, position = position_dodge(width = 0.1)) +
  geom_errorbar(aes(ymin = gm - sem, ymax = gm + sem), width = 0.1, position = position_dodge(width = 0.1)) +
  scale_color_manual(values = c("coral1", "blue")) +
  facet_wrap(vars(sex, treatment), ncol = 1, scales = "free") +
  theme_bw() +
  ylab("Geometric Mean Diameter (µm)") +
  theme(strip.text.x = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(angle = 90),
        legend.position = "top",
        axis.title.x = element_blank())
Size.plot
```



