---
title: "AST_histo_abundance_TP0"
author: "Chloé Gilligan"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Installing ggpubr package
```{r}
install.packages("ggpubr")
```

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("RColorBrewer")
library(ggpubr)
```

## Reading in data and metadata
```{r}
library(readr)
data <- read_csv("../data/astrangia_histology_abundance.csv", col_names = TRUE, na = c("", "NA")) %>%
select(full_sample_id, timepoint, treatment, date, species, month, site,  sex, StageI, StageII, StageIII, StageIV, StageV)
```


## Separating the "TP" from sample ID's the begin with AST, keeping them on the field samples
```{r}
library(stringr)
library(dplyr)
data <- data %>%
  mutate(full_sample_id = ifelse(startsWith(full_sample_id, "AST"), 
                                 gsub("-TP\\d+", "", full_sample_id), 
                                 full_sample_id))
data <- data %>%
  filter(!timepoint %in% c("TP2", "TP3", "TP4", "TP5", "TP6", "TP7"))

```

#calculate the total number of gametes in a sample
```{r}

data <- data %>%
rowwise() %>%
mutate(total.gametes = sum(StageI,StageII,StageIII,StageIV,StageV,na.rm = T))

```

# assign a value of Stage 0 to samples with no gametes
#if = if total.gametes = 0, then add a value of 1 to a new column called "Stage0"
#else = if total.gamates > 0, then add a value of 0 to a new column called "Stage0"
```{r}
data <- data %>%
  mutate(Stage0 = ifelse(total.gametes==0, 1, 0))
```

#calculate relative abundance of stage by sex, timepoint, treatment
```{r}
Rel.Abundance <- data %>%
  group_by(sex, timepoint, treatment, date) %>%
  summarise(
    total_gametes = sum(total.gametes, na.rm = TRUE),
    StageI = sum(StageI, na.rm = TRUE),
    StageII = sum(StageII, na.rm = TRUE),
    StageIII = sum(StageIII, na.rm = TRUE),
    StageIV = sum(StageIV, na.rm = TRUE),
    StageV = sum(StageV, na.rm = TRUE)
  ) %>%
  mutate(
    rel_abundance_StageI = StageI / total_gametes,
    rel_abundance_StageII = StageII / total_gametes,
    rel_abundance_StageIII = StageIII / total_gametes,
    rel_abundance_StageIV = StageIV / total_gametes,
    rel_abundance_StageV = StageV / total_gametes
  )

Rel.Abundance <- Rel.Abundance %>%
  filter(!sex=="NA") %>%
  filter(!treatment=="NA")
```

#Pivot longer for plotting
```{R}
Rel.Abundance_long <- Rel.Abundance %>%
  pivot_longer(cols = starts_with("rel_abundance_stage"), 
               names_to = "stage", 
               values_to = "value") %>%
  mutate(stage = factor(stage, 
levels = c("rel_abundance_StageI", "rel_abundance_StageII", "rel_abundance_StageIII", "rel_abundance_StageIV", "rel_abundance_StageV"), 
                        labels = c("Stage I", "Stage II", "Stage III", "Stage IV", "Stage V")))

```

# add lines of TP0 for both
#Creating a seperate dataset and plot for the field samples
```{r}
Rel.Abundance_long_field <- Rel.Abundance_long %>%
  filter(treatment=="Field")
 #mutate(date = as.Date(date, format = "%m/%d/%y"))  # Convert date column to Date type for proper ordering
```

# Convert the date column to Date type for proper ordering
```{r}
#Rel.Abundance_long_field <- Rel.Abundance_long_field %>%
  #mutate(date = as.Date(date, format = "%m/%d/%y"))
```

#Reordering dates
```{r}
# Reorder dates and relabel timepoints for field samples
Rel.Abundance_long_field <- Rel.Abundance_long %>%
  filter(treatment == "Field")%>%
mutate(date = factor(date, levels = c("4/11/21", "4/29/21", "6/4/21", "6/26/21", "7/24/21", "9/2/21"), labels = c("Early April", "Late April", "Early June", "Late June", "July", "September")))



```

```{r}
#Rel.Abundance_long_field <- Rel.Abundance_long_field %>%
 # mutate(month = case_when(
   # timepoint == "TP0" ~ "February",
   # timepoint == "TP1" ~ "March",
   # timepoint == "TP2" ~ "Early April",
   # timepoint == "TP3" ~ "Late April",
   # timepoint == "TP4" ~ "May",
    #timepoint == "TP5" ~ "June",
    #timepoint == "TP6" ~ "July",
    #timepoint == "TP7" ~ "August"
 # )) %>%
  #mutate(date = factor(date, levels = c("Early April", "Late April", "Early June", "Late June", "July", "September")))
```


#Plot for the realative proportion of Timepoint 0 Experiemental corals
```{r}
Rel.Abundance_long_tank <- Rel.Abundance_long %>%
  filter(!treatment=="Field")

ggplot(Rel.Abundance_long_tank, aes(x = date, y = value, fill = stage)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(sex ~ treatment) +
  theme_minimal() +
  labs(
    title = "Relative Abundance by Timepoint",
    x = "Timepoint",
    y = "Relative Abundance",
    color = "Stage")+
  labs(y = "Relative Proportion (%)", x = "Timepoint", title = "Relative Proportions of Timepoint 0 Experiemental Samples") +
  theme_bw() +
  scale_fill_manual(values = c("#CCEEF9", "#A6E1F4", "#59C7EB", "#00A9E0", "#137899"),
                    labels = c( 'Stage I', 'Stage II', 'Stage III', 'Stage IV', 'Stage V'))
```

